Confirmation of \srcfull{} as a long period radio transient

### 2025-05-15 paper/main.tex
Synch paper with Overleaf version

\srcfull{} (hereafter, \src{}) is a transient radio source discovered in the Variables and Slow Transients (VAST) survey \citep[][hereafter \citetalias{2024MNRAS.535..909D}]{2024MNRAS.535..909D}.
A single, highly polarized pulse was observed, lasting approximately 2 minutes, exhibiting a steep spectrum ($\alpha = -3.1$).
Extensive follow up in new and archival data at multiple wavelengths turned up no other pulses.
Despite this, the authors concluded that \src{} was most likely a long period transient (LPT), a newly emerging class of radio source with periods ranging from minutes to hours (see Table \ref{tbl:lpts} for an up-to-date list).

@@

[Reference for scattering kernels:] \citep{10.1093/mnras/stw1724}.

@@

>>>>

\srcfull{} (hereafter, \src{}) is a transient radio source discovered in the Variables and Slow Transients (VAST) survey \citep[][hereafter \citetalias{2024MNRAS.535..909D}]{2024MNRAS.535..909D}.
A single, highly polarized pulse was observed, lasting approximately 2 minutes, exhibiting a steep spectrum ($\alpha = -3.1$).
Extensive follow up in new and archival data at multiple wavelengths turned up no other pulses.
Despite this, the authors concluded that \src{} was most likely a long period transient (LPT), a newly emerging class of radio source with periods ranging from minutes to hours.

@@

%[Reference for scattering kernels:] \citep{10.1093/mnras/stw1724}.

@@

No significant polarised emission was detected in the MWA pulses, as well as the 2023 MeerKAT pulses.
The apparent lack of polarisation in the MeerKAT data set is likely due to low S/N coupled with the source not being not being centred in the beam.
Much more careful analysis would be needed to obtain reliable limits on the polarised components of those pulses.

The apparent lack of linear polarisation in the MWA pulses, however, is expected due to the high RM of ${\sim}961\,{\rm rad}/{\rm m}^2$ \citepalias[as reported in][]{2024MNRAS.535..909D}, causing the pulses to be depolarised within individual channels.
We also report that no significantly circularly polarised emission was seen at MWA frequencies.

The three pulses for which we report significant polarisation are (1) the 2023 ASKAP pulse \citepalias{2024MNRAS.535..909D}, (2) the 2024 ASKAP pulse, and (3) the 2024 MeerKAT pulse.
The high S/N of each time bin allowed us to test whether the rotation measure (RM) was consistent across each pulse.
We therefore performed a joint fit of the Stokes Q and U spectra of each time bin to
\begin{equation}
    Q(\nu) = {\rm Re}(\mathcal{L}(\nu)),
    \quad
    U(\nu) = {\rm Im}(\mathcal{L}(\nu)),
    \label{eqn:RM}
\end{equation}
where
\begin{equation}
    \mathcal{L}(\nu)
        = L_{\rm 1\,GHz}\left(\frac{\nu}{\rm 1\,GHz}\right)^{\alpha_L} e^{2i({\rm RM}\,\lambda^2 + \psi_0)},
\end{equation}
$L_{\rm 1\,GHz}$ is the flux density of the linear polarisation component at 1\,GHz, $\alpha_L$ is the spectral index of the linear component, RM is the rotation measure, $\lambda = c/\nu$ is the wavelength, and $\psi_0$ is the intrinsic polarisation angle.
The RMs and spectral indices of these fits are shown in the top two rows of panels of Fig. \ref{fig:RM}.

\begin{figure*}[th]
    \centering
    \includegraphics[width=0.98\linewidth]{RM.pdf}
    \caption{Polarisation of the 2023 and 2024 ASKAP pulses (left and middle) and the 2024 MeerKAT pulse (right). The top two rows show the RM and $\alpha_{\rm lin}$ parameters fitted to Eq. \eqref{eqn:RM} for each time bin independently. Only bins for which the uncertainty on $\alpha_{\rm lin}$ is less than $1.4$ are shown, a threshold which was found to cleanly eliminate off-pulse bins. The weighted average RMs for each pulse are indicated by the horizontal red lines. The PAs and lightcurves shown in the bottom two rows of panels were produced using only the average RM for the respective pulse.}
    \label{fig:RM}
\end{figure*}

In all cases, the fitted RM was within a few rad/m$^2$ of the value originally derived for the 2023 ASKAP pulse in \citetalias{2024MNRAS.535..909D}.
The RM appears to decrease slightly over the course of the 2023 ASKAP pulse, and \emph{increase} slightly over the course of the two 2024 MeerKAT components; however, all values are consistent with a constant RM across all three pulses.
The per-pulse weighted averaged values (with weighting proportional to the Stokes I profile) are $962 \pm 1$ (2023 ASKAP), $965 \pm 2$ (2024 ASKAP) and $962 \pm 2$ (2024 MeerKAT).
The spectral index of the linear component, $\alpha_L$, is similarly consistent, with averages value of $-2.1 \pm 0.4$ (2023 ASKAP), $-3.0 \pm 0.6$ (2024 ASKAP) and $-1.9 \pm 0.5$ (2024 MeerKAT).

Each pulse was de-Faraday rotated with the weighted average RM given above, to derive the linear polarisation profile and the PA evolution across the pulses.
For the 2023 ASKAP pulse, we recover the `S'-shaped PA curve discussed at length in \citetalias{2024MNRAS.535..909D} in the context of the RVM.
Contrastingly, the PA curve of the 2024 ASKAP pulse is flat (across five 10-second bins), and that of the 2024 MeerKAT pulse is curved, but not in an RVM-like way.
